Some examples in extended logic counting classes

### 2016-12-05 LICS/appendix.tex
standarizing saluja

@@

@@

Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
\begin{claim} \label{mono}
	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
\end{claim}
\begin{proof}
	($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$.

@@

Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there exists $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
\begin{eqnarray*}
	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
\end{eqnarray*}
Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
\begin{multline*}
\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
\end{multline*}
Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (T(x) \vee \neg T(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. Given that $f_{\#3DNF}, f_{all} \in \E1$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E1$. Let $h\in\E1$ be such that $h = f_{all}-f_{\#3DNF}$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete. However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E1$, we have that $L_h \in \ptime$. Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.

@@

We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.

@@

Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
	\begin{eqnarray*}
		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
	\end{eqnarray*}
	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
	\begin{eqnarray*}
		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
	\end{eqnarray*}
	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$

	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
	\begin{eqnarray}
	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
	\end{eqnarray}
	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
	\begin{multline*}
	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
	\end{multline*}
	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$. Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.

@@

	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$

	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$. In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
	\begin{eqnarray*}
		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
	\end{eqnarray*}
	and
	\begin{eqnarray*}
		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
	\end{eqnarray*}
	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables. Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2. The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.

>>>>

\textbf{Proof of Theorem \ref{no-mult}}

Let $\{+, \cdot, \Sigma, \Pi\}$ be the set of {\it algebraic operators} in $\qso$. We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition. Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$. The case $\text{op}(\alpha) = 0$ is trivial. Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition. Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$. Note that if $\alpha$ is of the form $\alpha = (\beta_1 + \beta_2)$, $\alpha = \Sigma x \beta$, $\alpha = \Pi x \beta$, $\alpha = \Sigma X \beta$ or $\alpha = \Pi X \beta$ the proof follows directly. Therefore we assume that $\alpha = (\beta_1 \cdot \beta_2)$. We separate the proof in two main cases:
\begin{enumerate}
	\item At least one of $\beta_1, \beta_2$ is not of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
	\begin{enumerate}
		\item $\beta_1 = s \in \nat$. Then we use $\alpha' = \beta_2 + \cdots + \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$. Since each of the formulas paired by $+$ has less than $n$ algebraic operators, the statement holds.
		\item $\beta_1 = (\gamma_1 + \gamma_2)$. Then, we use $\alpha' = (\gamma_1\cdot\beta_2 + \gamma_2\cdot\beta_2)$ and each of the formulas paired by $+$ have less than $n$ algebraic operators so the statement holds.
		\item $\beta_1 = \Sigma x \gamma$. Then, we use $\alpha' = \Sigma x (\gamma\cdot\beta_2)$ and as in the previous case the statement holds.
		\item $\beta_1 = \Sigma X \gamma$, which we treat exactly as the previous one.
	\end{enumerate}
	\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases. We define $(\varphi \mapsto \alpha) := (\neg\varphi + (\varphi\cdot\alpha))$ for which $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise.
	\begin{enumerate}
		\item $\beta_1 = \Sigma x \gamma_1(x)$ and $\beta_2 = \Sigma y \gamma_2(y)$. Then we use $\alpha' = \Sigma x(\gamma_1(x)\cdot\gamma_2(x))$, and the statement holds.
		\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma Y \gamma_2(Y)$. [something about arities].
		\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma x \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
		\item $\beta_1 = \Sigma x \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \Sigma x (\gamma \cdot \varphi)$, and the statement holds.
		\item $\beta_1 = \Sigma X \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
	\end{enumerate}
\end{enumerate}

@@

\textbf{Proof of Theorem \ref{saluja-eq}}

@@

Let $f$ be a function in $\E{1}$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
\begin{claim} \label{mono}
	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
\end{claim}
\begin{proof}
	($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$.

@@

Suppose that $\E{1}$ is closed under substraction, that is, for each pair of functions $f,g\in \E{1}$, there exists $h\in\E{1}$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
\begin{eqnarray*}
	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
\end{eqnarray*}
Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
\begin{multline*}
\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
\end{multline*}
Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (T(x) \vee \neg T(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. Given that $f_{\#3DNF}, f_{all} \in \E{1}$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E{1}$. Let $h\in\E{1}$ be such that $h = f_{all}-f_{\#3DNF}$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete. However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E{1}$, we have that $L_h \in \ptime$. Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.

@@

We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.

@@

Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
	\begin{eqnarray*}
		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
	\end{eqnarray*}
	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
	\begin{eqnarray*}
		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
	\end{eqnarray*}
	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$

	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
	\begin{eqnarray}
	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
	\end{eqnarray}
	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
	\begin{multline*}
	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
	\end{multline*}
	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$. Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.

@@

	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$

	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$. In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
	\begin{eqnarray*}
		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
	\end{eqnarray*}
	and
	\begin{eqnarray*}
		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
	\end{eqnarray*}
	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables. Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2. The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
